\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing income inequality with tax data - Switzerland from 1917 to 2010}

### 2014-07-04 paper/subdocuments/data.Rnw
textbf -> textit

\emph{Measuring inequality or concentration} \\
To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension. The Gini coefficient is the most known measure and mainly used for international comparison. As it is derived from the Lorenz-curve, the quantified amount of inequality can unpretentiously be described in a formal and visual way. Therefore the Gini coefficient is easily interpretable. Furthermore it has several desired statistical properties \citet{engelhardt_modelle_2000}. (1) ``\textbf{principle of population}'': the assessment of inequality is independent of the population size (2) ``\textbf{Requirement of Bresciani-Turroni}'': the measure is sensitive for changes of income shares, but not for absolute changes (e.g. doubling of all income) (3) ``\textit{weak principle of transfers}'' or ``\textit{requirement of Pigou-Dalton}'': transfers from richer households lead to a reduction of inequality. However, several drawbacks are reported in the literature. The most important point is, that the underlying distributional form of the measured inequality is unknown and it is therefore not possible to see if the measure is driven by a few rich or many poor individuals. This can also be problematic for comparison between countries or over time. In extreme cases two totally different distributions share the same Gini-coefficient.\\

>>>>

\emph{Measuring inequality or concentration} \\
To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension. The Gini coefficient is the most known measure and mainly used for international comparison. As it is derived from the Lorenz-curve, the quantified amount of inequality can unpretentiously be described in a formal and visual way. Therefore the Gini coefficient is easily interpretable. Furthermore it has several desired statistical properties \citet{engelhardt_modelle_2000}. (1) ``\textit{principle of population}'': the assessment of inequality is independent of the population size (2) ``\textit{Requirement of Bresciani-Turroni}'': the measure is sensitive for changes of income shares, but not for absolute changes (e.g. doubling of all income) (3) ``\textit{weak principle of transfers}'' or ``\textit{requirement of Pigou-Dalton}'': transfers from richer households lead to a reduction of inequality. However, several drawbacks are reported in the literature. The most important point is, that the underlying distributional form of the measured inequality is unknown and it is therefore not possible to see if the measure is driven by a few rich or many poor individuals. This can also be problematic for comparison between countries or over time. In extreme cases two totally different distributions share the same Gini-coefficient.\\

### 2014-07-04 paper/subdocuments/results.Rnw
textbf -> textit

<<with_without_zeros_Gini,warning=FALSE, fig.cap='Gini coefficients with, without and imputed non-taxed units',fig.pos='H'>>=
df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
#library(reshape)
dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
names(dflong)[3]<-"Type"names(dflong)[4]<-"Gini"
levels(dflong$Type)<-c("without zeros","including zeros","imputed values for zeros")
ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], 
       aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+
  scale_y_continuous(limits=c(0.18,0.6))+
  scale_x_continuous(limits=c(1994,2012),breaks=number_ticks(10)) +
  geom_line(size=0.5)+
  geom_point(aes(shape=Type),size=3)+
  theme_bw()+
  xlab("tax period")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))

@@

estv_gini <- estv %.% 
  filter(kanton=="CH",eink_art=="Nach steuerbarem Einkommen",fall=="Normalfälle",anz_pflichtige>0,steuerperiode!=1945.5) %.% 
  select(bemessungsjahre, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %.%
  group_by(factor(bemessungsjahre)) %.%
  mutate(mean_steink=steink/anz_pflichtige) %.%
  summarise(gini(mean_steink,weights=anz_pflichtige))

>>>>

<<with_without_zeros_Gini,warning=FALSE, fig.cap='Gini coefficients with, without and imputed non-taxed units',fig.pos='H'>>=
df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
#library(reshape)
dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
names(dflong)[3]<-"Type"
names(dflong)[4]<-"Gini"
levels(dflong$Type)<-c("without zeros","including zeros","imputed values for zeros")
ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], 
       aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+
  scale_y_continuous(limits=c(0.18,0.6))+
  scale_x_continuous(limits=c(1994,2012),breaks=number_ticks(10)) +
  geom_line(size=0.5)+
  geom_point(aes(shape=Type),size=3)+
  theme_bw()+
  xlab("tax period")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))

@@

#estv_gini <- estv %.% 
#  filter(kanton=="CH",eink_art=="Nach steuerbarem Einkommen",fall=="Normalfälle",anz_pflichtige>0,steuerperiode!=1945.5) %.% 
#  select(bemessungsjahre, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %.%
#  group_by(factor(bemessungsjahre)) %.%
#  mutate(mean_steink=steink/anz_pflichtige) %.%
#  summarise(gini(mean_steink,weights=anz_pflichtige))

load("estv_gini.RData")
